---
title: "pres-script"
author: "Sebastian Wagner"
date: "2023-06-08"
output: pdf_document
---

```{r hidden=TRUE}
# for pretty printing
library(glue)

```

## Data Preparation

```{r}
# set working dir to where data is
#setwd("INSERT")


# Imported preprocessed data
dat <- read.csv("Data_clean/log_p6.txt", sep="|")

# Filter sitting data out
sitting_phone <- dat[dat$statusId == 2, ] # this is the id for sitting on phone
sitting_no_phone <- dat[dat$statusId == 1, ] # this is the id for sitting
walking_phone <- dat[dat$statusId == 4, ] # this is the id for walking on phone
walking_no_phone <- dat[dat$statusId == 3, ] # this is the id for walking


## Timestamp = milliseconds after experiment start
mt_1 <- min(sitting_phone$timestamp)
sitting_phone$timestamp <- sitting_phone$timestamp - mt_1

mt_2 <- min(sitting_no_phone$timestamp)
sitting_no_phone$timestamp <- sitting_no_phone$timestamp - mt_2

mt_3 <- min(walking_phone$timestamp)
walking_phone$timestamp <- walking_phone$timestamp - mt_3

mt_4 <- min(walking_no_phone$timestamp)
walking_no_phone$timestamp <- walking_no_phone$timestamp - mt_4

# Acceleration level can be calculated with this:
sitting_phone$acceleration <- sqrt(sitting_phone$x^2 + sitting_phone$y^2 + sitting_phone$z^2)
sitting_no_phone$acceleration <- sqrt(sitting_no_phone$x^2 + sitting_no_phone$y^2 + sitting_no_phone$z^2)
walking_phone$acceleration <- sqrt(walking_phone$x^2 + walking_phone$y^2 + walking_phone$z^2)
walking_no_phone$acceleration <- sqrt(walking_no_phone$x^2 + walking_no_phone$y^2 + walking_no_phone$z^2)

sitting_phone <- sitting_phone[c("timestamp", "acceleration")]
sitting_no_phone <- sitting_no_phone[c("timestamp", "acceleration")]
walking_phone <- walking_phone[c("timestamp", "acceleration")]
walking_no_phone <- walking_no_phone[c("timestamp", "acceleration")]

# Calculate the common y-axis limits
y_limits <- c(min(sitting_phone$acceleration, walking_phone$acceleration, sitting_no_phone$acceleration, walking_no_phone$acceleration), 
max(sitting_phone$acceleration, walking_phone$acceleration, sitting_no_phone$acceleration, walking_no_phone$acceleration))

# Calculate the common x-axis limits
x_limits <- c(min(sitting_phone$timestamp, walking_phone$timestamp, sitting_no_phone$timestamp, walking_no_phone$timestamp),
max(sitting_phone$timestamp, walking_phone$timestamp, sitting_no_phone$timestamp, walking_no_phone$timestamp))

# Plot on phone data
plot(sitting_phone$timestamp,sitting_phone$acceleration,type = "l",col = 2,xlab = "Timestamp",ylab = "Acceleration sitting on phone", ylim = y_limits, xlim = x_limits)
plot(walking_phone$timestamp,walking_phone$acceleration,type = "l",col = 4,xlab = "Timestamp",ylab = "Acceleration walking on phone", ylim = y_limits, xlim = x_limits)

# Plot off phone data
plot(sitting_no_phone$timestamp,sitting_no_phone$acceleration,type = "l",col = 2,xlab = "Timestamp",ylab = "Acceleration walking not on phone", ylim = y_limits, xlim = x_limits)
plot(walking_no_phone$timestamp,walking_no_phone$acceleration,type = "l",col = 4,xlab = "Timestamp",ylab = "Acceleration sitting not on phone", ylim = y_limits)

# We will minimize the effects of starting phase on the statistical analysis by
# eliminating the first 3000 timestamps:
sitting_phone <- sitting_phone[-(1:3000),]
sitting_no_phone <- sitting_no_phone[-(1:3000),]
walking_phone <- walking_phone[-(1:3000),]
walking_no_phone <- walking_no_phone[-(1:3000),]

# Define means
mean_sitting_phone <- mean(sitting_phone$acceleration)
mean_sitting_no_phone <- mean(sitting_no_phone$acceleration)
mean_walking_phone <- mean(walking_phone$acceleration)
mean_walking_no_phone <- mean(walking_no_phone$acceleration)

# Create a data frame from the means
means_dataset <- data.frame(
  Condition = c("Sitting with Phone", "Sitting without Phone", "Walking with Phone", "Walking without Phone"),
  Mean_Acceleration = c(mean_sitting_phone, mean_sitting_no_phone, mean_walking_phone, mean_walking_no_phone)
)

# Print the dataset
means_dataset

```

## Testing

> We expect to observe lower acceleration levels during sitting while using the phone compared to sitting without phone usage.
> We expect to observe higher acceleration levels during walking while using the phone compared to walking without phone usage.

```{r}
# Plot on phone data
plot(sitting_phone$timestamp,sitting_phone$acceleration,type = "l",col = 2,xlab = "Timestamp",ylab = "Acceleration")

# Plot off phone data
plot(sitting_no_phone$timestamp,sitting_no_phone$acceleration,type = "l",col = 2,xlab = "Timestamp",ylab = "Acceleration no phone")

# Plot acceleration frequency
hist(sitting_phone$acceleration, breaks=500)
hist(sitting_no_phone$acceleration, breaks=500)
hist(walking_phone$acceleration, breaks=500)
hist(walking_no_phone$acceleration, breaks=500)

# We can see that our data does not necessarily follow a normal distribution
# This means we use Mann-Whitney test
wilcox.test(sitting_phone$acceleration, sitting_no_phone$acceleration)
# Sitting while not on the phone has less acceleration

mean(sitting_phone$acceleration)
mean(sitting_no_phone$acceleration)

wilcox.test(walking_phone$acceleration, walking_no_phone$acceleration)
# Walking while not on the phone has less acceleration

median(walking_phone$acceleration)
median(walking_no_phone$acceleration)

# Let's see what the p value of testing the same class would be
wilcox.test(walking_phone$acceleration[1:15000], walking_phone$acceleration[15000:30000])
```

## Multiple Testing

```{r}
# We've cramped the procedure in this function:
test_participant <- function(filename) {
  dat <- read.csv(filename, sep="|")
  sitting_phone <- dat[dat$statusId == 2, ] # this is the id for sitting on phone
  sitting_no_phone <- dat[dat$statusId == 1, ] # this is the id for sitting
  walking_phone <- dat[dat$statusId == 4, ] # this is the id for walking on phone
  walking_no_phone <- dat[dat$statusId == 3, ] # this is the id for walking
  sitting_phone <- sitting_phone[-(1:3000),]
  sitting_no_phone <- sitting_no_phone[-(1:3000),]
  walking_phone <- walking_phone[-(1:3000),]
  walking_no_phone <- walking_no_phone[-(1:3000),]
  sitting_phone$acceleration <- sqrt(sitting_phone$x^2 + sitting_phone$y^2 + sitting_phone$z^2)
  sitting_no_phone$acceleration <- sqrt(sitting_no_phone$x^2 + sitting_no_phone$y^2 + sitting_no_phone$z^2)
  walking_phone$acceleration <- sqrt(walking_phone$x^2 + walking_phone$y^2 + walking_phone$z^2)
  walking_no_phone$acceleration <- sqrt(walking_no_phone$x^2 + walking_no_phone$y^2 + walking_no_phone$z^2)
  # Print p-value of test
  sitting_res = glue('Sitting p = {wilcox.test(sitting_phone$acceleration, sitting_no_phone$acceleration)$p.value}')
  walking_res = glue('Walking p = {wilcox.test(walking_phone$acceleration, walking_no_phone$acceleration)$p.value}')
  
  glue('{sitting_res}\n{walking_res}')
}

# Test all participants
test_participant("Data_clean/log_p1.txt")
test_participant("Data_clean/log_p2.txt")
test_participant("Data_clean/log_p3.txt")
test_participant("Data_clean/log_p4.txt")
test_participant("Data_clean/log_p5.txt")
test_participant("Data_clean/log_p6.txt")
test_participant("Data_clean/log_p7.txt")
test_participant("Data_clean/log_p8.txt")
test_participant("Data_clean/log_p9.txt")
test_participant("Data_clean/log_p10.txt")

# All of the p values are very low and fall below 5 percent threshold
# Even when you corrent it with Bonferroni correction the threshold for each 
# test is alpha / 10 = 0.005 which they are still below

```


Now, for the feasibility study we have 10 recordings from different people that we'll use.
In the following we will perform a test for all of the data combined.
We'll be doing the above procedure for each participants data.
In the end this will amount to 10 * 3 = 30 F-tests for each participant and dimension.
For the test we will arbitrarily set our alpha to 5% and correct it with the Bonferroni correction.
We divide it by the number of tests performed, to get an acceptable alpha of 0.05 / 30 = 0.00167 per test.

